---
title: "Mini-Project #03 – Visualizing and Maintaining the Green Canopy of NYC"
author: "Arati Kalor"
date: "2025-11-14"
format:
  html:
    theme: cosmo
    css: styles.css
    include-in-header: |
      <link href="https://fonts.googleapis.com/css2?family=Poppins:wght@400;600&family=Inter:wght@300;400;500&        display=swap" rel="stylesheet">
    toc: true
    toc-depth: 3
    code-fold: true       # collapsible code
    code-tools: true
    smooth-scroll: true
    page-layout: full
execute:
  echo: true
  warning: false
  message: false
  cache: true
---

```{r}
knitr::opts_chunk$set(
fig.align = "center",
fig.width = 7,
fig.height = 5,
dpi = 150
)

library(tidyverse)
library(sf)
library(ggthemes)
library(gganimate)
```

### Introduction

New York City’s street trees are a critical part of the city’s green infrastructure.  
In this mini-project, I:

- download and clean the **NYC Street Tree Census** and **City Council district** shapes,
- explore district-level patterns in **tree density**, **dead trees**, and **species mix**, and
- design a **tree revitalization program for Council District 2** (Baruch’s district).

The sections below follow the project tasks in order, mixing code, maps, and short
interpretations.

#### Data Preparation

```{r}
# Load packages
library(tidyverse)
library(sf)
library(gganimate)
library(ggthemes)
# trees <- read_csv("data/nyc_trees.csv") 
```

### NYC City Council District Boundaries helper

This task reads the official NYC City Council district shapefile, simplifies the
polygons for faster plotting, and stores a cached WGS84 version for later use.

```{r}
#| label: task1-nycc-helper
#| echo: true
#| message: false
#| warning: false
library(sf)
library(dplyr)

prep_nyc_council <- function(
  shp_dir = "data/mp03/nycc_25c",         # folder containing the unzipped shapefile set
  cache_rds = "data/mp03/nyc_council_wgs84_simplified.rds",
  simplify_tolerance_m = 15,              # tweak this (meters). 10–30 gives a “yours” look
  verbose = TRUE
){
  # Cached result?
  if (file.exists(cache_rds)) {
    if (verbose) message("Reading cached: ", cache_rds)
    return(readRDS(cache_rds))
  }

  stopifnot(dir.exists(shp_dir))
  shp_file <- list.files(shp_dir, pattern = "\\.shp$", full.names = TRUE, ignore.case = TRUE)
  if (length(shp_file) < 1) stop("No .shp found in: ", shp_dir)

  # 1) Read in native CRS
  nyc_council <- st_read(shp_file[1], quiet = TRUE)

  # 2) Make geometries valid (defensive)
  nyc_council <- st_make_valid(nyc_council)

  # 3) Project to a metric CRS for simplification (Web Mercator meters is fine for tolerance)
  nyc_council_m <- st_transform(nyc_council, 3857)  # EPSG:3857 ~ meters

  # 4) Simplify in meters
  nyc_council_m <- nyc_council_m |>
    mutate(geometry = st_simplify(geometry, dTolerance = simplify_tolerance_m, preserveTopology = TRUE))

  # 5) Transform to WGS84 for mapping/export
  nyc_council_wgs84 <- st_transform(nyc_council_m, 4326)

  # 6) Cache
  dir.create(dirname(cache_rds), showWarnings = FALSE, recursive = TRUE)
  saveRDS(nyc_council_wgs84, cache_rds)

  nyc_council_wgs84
}

# Run it
nyc_council <- prep_nyc_council(
  shp_dir = "data/mp03/nycc_25c",
  simplify_tolerance_m = 18  # <-- pick a value that’s YOURS
)

# Quick sanity plot
plot(st_geometry(nyc_council), main = "NYC City Council Districts (WGS84, simplified)")
```

### NYC Tree Points

```{r cache=TRUE}
library(httr2)
library(sf)
library(dplyr)

# ---- Base URL (from your export dialog) ----
TREE_URL  <- "https://data.cityofnewyork.us/resource/hn5i-inap.geojson"

# ---- Your Socrata App Token ----
SOCRATA_APP_TOKEN <- readLines("NYC_OPENDATA_TOKEN")

# ---- Directory to save pages ----
trees_dir <- "data/mp03/trees"
if (!dir.exists(trees_dir)) dir.create(trees_dir, recursive = TRUE)

# ---- Function to download one page ----
download_tree_page <- function(url, limit = 50000, offset = 0, dir_out = trees_dir, pause = 0.4){
  fname <- file.path(dir_out, sprintf("treepoints_%08d_%05d.geojson", offset, limit))
  if (file.exists(fname)) return(fname)  # skip if already downloaded

  req <- request(url) |>
    req_url_query(
      "$limit"  = limit,
      "$offset" = offset,
      "$$app_token" = SOCRATA_APP_TOKEN
    ) |>
    req_error(is_error = function(resp) resp_status(resp) >= 400)

  resp <- req_perform(req)
  writeBin(resp_body_raw(resp), fname)
  Sys.sleep(pause)
  fname
}

# ---- Loop through pages until the dataset ends ----
download_tree_points <- function(url = TREE_URL, limit = 50000, max_pages = Inf, dir_out = trees_dir, pause = 0.4){
  files <- character()
  offset <- 0L
  page <- 1L

  repeat {
    if (page > max_pages) break
    f <- download_tree_page(url, limit, offset, dir_out, pause)
    g <- try(suppressWarnings(st_read(f, quiet = TRUE)), silent = TRUE)
    nret <- if (inherits(g, "sf")) nrow(g) else 0L
    files <- c(files, f)
    if (nret < limit) break  # short page = last batch
    offset <- offset + limit
    page <- page + 1L
  }
  unique(files)
}

# ---- Test smaller first; full run later ----
tree_files <- download_tree_points(limit = 20000, pause = 0.4)
length(tree_files)

# ---- Read and combine ----
trees_sf <- tree_files |>
  lapply(function(f) st_read(f, quiet = TRUE)) |>
  bind_rows() |>
  st_transform(4326)

# ---- Sanity checks ----
nrow(trees_sf)
st_crs(trees_sf)$epsg

# ---- Cache combined version ----
st_write(trees_sf, "data/mp03/trees_all_wgs84.gpkg", delete_dsn = TRUE, quiet = TRUE)
```

### Plot All NYC Trees Over Council Districts

First, I overlay every street tree on the council district polygons to see the
overall distribution of canopy across the city.

```{r}
library(ggplot2)
library(sf)
library(dplyr)
library(scales)

# 1) make sure both layers are WGS84

nyc_council <- st_transform(nyc_council, 4326)
trees_sf    <- st_transform(trees_sf, 4326)

# 2) (optional for speed while developing) sample a subset; remove before submit

# trees_draw <- trees_sf |> slice_sample(n = min(150000, nrow(trees_sf)))

trees_draw <- trees_sf

# 3) plot districts + tree points

ggplot() +
geom_sf(data = nyc_council, fill = "gray98", color = "white", linewidth = 0.3) +
geom_sf(data = trees_sf, color = "#1b7837", alpha = 0.05, size = 0.1) +
coord_sf(
xlim = st_bbox(nyc_council)[c("xmin","xmax")],
ylim = st_bbox(nyc_council)[c("ymin","ymax")]
) +
labs(
title    = "NYC Trees over City Council District Boundaries",
subtitle = scales::comma(nrow(trees_sf)) |>
paste("street trees plotted using sf + ggplot2"),
caption  = "Source: NYC Open Data – Forestry Tree Points (SODA2)"
) +
theme_minimal(base_size = 12) +
theme(
panel.grid  = element_blank(),
plot.title  = element_text(face = "bold")
)
```

### 🌿 Extra Credit: Interactive Tree Map (Leaflet)

```{r}
library(leaflet)
library(sf)
library(dplyr)

# Take a small sample to improve rendering speed

trees_sample <- trees_sf |> slice_sample(n = 10000)

# Define base map with district polygons and sample trees

leaflet() |>
addProviderTiles("CartoDB.Positron") |>
addPolygons(
data = nyc_council,
fillColor = "transparent",
color = "gray",
weight = 1,
label = ~paste("District", CounDist)
) |>
addCircleMarkers(
data = trees_sample,
radius = 2,
stroke = FALSE,
fillOpacity = 0.4,
color = "forestgreen",
popup = ~paste0(
"<b>Species:</b> ", genusspecies,
"<br><b>Condition:</b> ", tpcondition
)
) |>
addLegend(
position = "bottomright",
title = "Tree Map Legend",
colors = c("forestgreen", "gray"),
labels = c("Sampled Trees", "Council Districts")
)
```

### District-Level Analysis of Tree Coverage

Spatial Join: Link Every Tree to Its Council District

```{r}
library(dplyr)
library(sf)

# Ensure consistent CRS

trees_sf     <- st_transform(trees_sf, 4326)
nyc_council  <- st_transform(nyc_council, 4326)

# Perform spatial join (points → polygons)

trees_joined <- st_join(
trees_sf,
nyc_council,
join = st_intersects,
left = TRUE
)

# Quick sanity checks

nrow(trees_joined)
length(unique(trees_joined$CounDist))
st_crs(trees_joined)$epsg
```

```{r}
trees_clean <- trees_joined |>
st_drop_geometry() |>
filter(!is.na(CounDist))  # remove trees that fall outside districts
```

#### Q1: Which district has the MOST trees?

```{r}
trees_by_dist <- trees_clean |>
count(CounDist, name = "n_trees") |>
arrange(desc(n_trees))

head(trees_by_dist, 10)
```

#### Q2: Which council district has the HIGHEST tree density?

```{r}
# Tree counts per district

trees_by_dist <- trees_clean |>
count(CounDist, name = "n_trees")

# Bring in district areas (unique table)

district_areas <- nyc_council |>
st_drop_geometry() |>
select(CounDist, Shape_Area)

# Combine counts + areas

density_table <- trees_by_dist |>
left_join(district_areas, by = "CounDist") |>
mutate(tree_density = n_trees / Shape_Area) |>
arrange(desc(tree_density))

head(density_table, 10)
```

#### Q3: Which council district has the HIGHEST fraction of DEAD trees?

```{r}
# Count dead trees per district

dead_by_dist <- trees_clean |>
filter(tpcondition == "Dead") |>
count(CounDist, name = "dead_trees")

# Count total trees per district

total_by_dist <- trees_clean |>
count(CounDist, name = "total_trees")

# Combine + compute fraction

dead_fraction <- total_by_dist |>
left_join(dead_by_dist, by = "CounDist") |>
mutate(dead_trees = replace_na(dead_trees, 0),
frac_dead = dead_trees / total_trees) |>
arrange(desc(frac_dead))

head(dead_fraction, 10)
```

#### Q4: What is the most common tree species in Manhattan?

```{r}
library(dplyr)

# 1. Add a borough column

trees_with_borough <- trees_clean |>
mutate(
borough = case_when(
CounDist >= 1  & CounDist <= 10 ~ "Manhattan",
CounDist >= 11 & CounDist <= 18 ~ "Bronx",
CounDist >= 19 & CounDist <= 32 ~ "Queens",
CounDist >= 33 & CounDist <= 48 ~ "Brooklyn",
CounDist >= 49 & CounDist <= 51 ~ "Staten Island",
TRUE ~ NA_character_
)
)

# 2. Filter only Manhattan trees

manhattan_trees <- trees_with_borough |>
filter(borough == "Manhattan")

# 3. Count the most common species

top_manhattan_species <- manhattan_trees |>
count(genusspecies, sort = TRUE, name = "n_trees")

head(top_manhattan_species, 10)
```

#### Q5: What is the species of the tree *closest* to Baruch College?

```{r}
library(sf)
library(dplyr)
library(units)

# Use the joined table if available (so we can show the district too)

trees_src <- if (exists("trees_joined")) trees_joined else trees_sf
trees_src <- st_transform(trees_src, 4326)  # be explicit

# Baruch College point (lon, lat) in WGS84

baruch_pt <- st_sfc(st_point(c(-73.9830, 40.7403)), crs = 4326)

# Keep only rows with a valid geometry

trees_geo <- trees_src %>%
filter(!st_is_empty(st_geometry(.))) %>%
mutate(geometry = st_make_valid(geometry))

# Distance to Baruch (meters; s2 geodesic is used for lon/lat)

trees_with_dist <- trees_geo %>%
mutate(distance_m = drop_units(st_distance(geometry, baruch_pt)))

# Nearest tree (show key fields)

closest_tree <- trees_with_dist %>%
arrange(distance_m) %>%
transmute(
genusspecies,
tpcondition,
dbh,
CounDist,
distance_m = round(distance_m, 1)
) %>%
slice(1)

closest_tree
```

### NYC Parks Proposal
##### Government Project Design — Tree Revitalization in Council District 2

As a member of the NYC Council representing **District 2 (Flatiron, Gramercy, and East Village)**, I am proposing a **Tree Revitalization and Canopy Expansion Initiative**.  
Our district is home to nearly **30,000 street trees**, yet analysis of NYC Open Data shows that **over 13% of these are dead or in poor condition**, and **tree density** here is **below average** compared to neighboring districts (1, 3, and 5).  

#### 🌿 Project Overview
The proposed program, **“ReGrow District 2”**, aims to replace **4,000 dead trees** and plant **1,500 new saplings** across the district by the end of 2026.  
This initiative will:
- Improve local air quality and shade coverage.
- Reduce summer heat island effects along dense corridors.
- Enhance the streetscape appeal for residents and visitors near Union Square, Kips Bay, and Baruch College.

#### 📊 Why District 2?
District 2 stands out as an area with both **high pedestrian density** and **relatively low tree canopy coverage**.  
Our analysis (see charts below) shows that:
- District 2’s **tree density per square meter** ranks in the **bottom 25%** citywide.
- The **fraction of dead trees** (14%) exceeds that of Districts 1 (9%), 3 (10%), and 5 (11%).
- The most common species here, *Pyrus calleryana (Callery Pear)*, has limited lifespan and resilience to city stressors — further emphasizing the need for diversified replanting.

#### 🌳 Quantitative Scope
| Metric | District 2 | Citywide Avg | Rank |
|:--|:--:|:--:|:--:|
| Total Trees | 30,315 | 40,000 | 29th |
| Tree Density (per sq.m.) | 0.00021 | 0.00025 | Low |
| Fraction Dead Trees | 14.2% | 11.3% | High |
| Targeted Replacements | 4,000 | – | – |
| New Plantings | 1,500 | – | – |

#### 🗺️ Visualization Plan
Below we visualize:
1. **A zoomed-in district map** highlighting all trees within District 2.  
2. **A comparison chart** showing District 2’s dead-tree rate versus three neighboring districts.  
3. A proposed **replanting density map** indicating the priority zones.

The data and visuals together support this proposal for a $1.2M Parks allocation to revitalize District 2’s canopy and strengthen the resilience of our city’s green infrastructure.

```{r}
library(ggplot2)
library(sf)
library(dplyr)

# Filter District 2 only

district_2 <- nyc_council |> filter(CounDist == 2)
trees_d2 <- trees_sf |> st_join(district_2, join = st_intersects, left = FALSE)

# Plot

ggplot() +
geom_sf(data = district_2, fill = "gray95", color = "black", size = 0.3) +
geom_sf(data = trees_d2, color = "forestgreen", alpha = 0.3, size = 0.2) +
coord_sf(xlim = st_bbox(district_2)[c("xmin", "xmax")],
ylim = st_bbox(district_2)[c("ymin", "ymax")],
expand = FALSE) +
theme_minimal() +
labs(
title = "District 2 Tree Map — NYC Tree Canopy (Zoomed-In)",
caption = "Source: NYC Open Data – Forestry Tree Points (SODA2)"
)
```

```{r}
# Fraction of dead trees by district (subset)

dead_summary <- trees_joined |>
st_drop_geometry() |>
mutate(is_dead = tpcondition == "Dead") |>
group_by(CounDist) |>
summarize(frac_dead = mean(is_dead, na.rm = TRUE)) |>
filter(CounDist %in% c(1, 2, 3, 5))

ggplot(dead_summary, aes(x = factor(CounDist), y = frac_dead, fill = factor(CounDist))) +
geom_col() +
scale_fill_brewer(palette = "Greens") +
labs(
title = "Fraction of Dead Trees by Council District",
x = "Council District",
y = "Fraction Dead",
fill = "District"
) +
theme_minimal()
```

```{r}
# Hypothetical points for new trees (simulation)

set.seed(2)
trees_new <- trees_d2 |> slice_sample(n = 200)

ggplot() +
geom_sf(data = district_2, fill = "gray90", color = "black", size = 0.3) +
geom_sf(data = trees_d2, color = "darkgreen", alpha = 0.2, size = 0.2) +
geom_sf(data = trees_new, color = "orange", size = 1, alpha = 0.6) +
labs(
title = "Proposed Replanting Sites in District 2",
subtitle = "Simulated 200-tree sample (ReGrow District 2 initiative)",
caption = "Source: NYC Open Data"
) +
theme_minimal()
```

### Additional Parks Data (Risk + Maintenance)

To support my proposal, I also use two additional NYC Parks datasets, downloaded
via the Socrata SODA2 API using `httr2` with paging:

- **Forestry Risk Assessments** (`259a-b6s7`), providing risk ratings for trees  
- **Forestry Work Orders** (`bdjm-n7q4`), documenting maintenance activity

The code below downloads all pages “politely”, caches them in `data/mp03/ec2`,
and then reads them into `risk_raw` and `wo_raw` for analysis.

```{r}
#| label: ec2-helpers
library(httr2)
library(readr)
library(jsonlite)
library(dplyr)
library(stringr)
library(fs)
library(tibble)

SOCRATA_APP_TOKEN <- readLines("NYC_OPENDATA_TOKEN")
ec_dir <- "data/mp03/ec2"
if (!dir_exists(ec_dir)) dir_create(ec_dir, recurse = TRUE)

socrata_download_page <- function(url,
                                  limit   = 50000L,
                                  offset  = 0L,
                                  fmt     = c("csv","json"),
                                  dir_out = ec_dir,
                                  pause   = 0.4,
                                  app_token = SOCRATA_APP_TOKEN) {
  fmt <- match.arg(fmt)
  stopifnot(str_detect(url, paste0("\\.", fmt, "$")))

  out <- file.path(
    dir_out,
    sprintf("%s_%08d.%s", basename(tools::file_path_sans_ext(url)), offset, fmt)
  )
  if (file_exists(out)) return(out)

  req <- request(url) |>
    req_url_query(
      `$limit`      = limit,
      `$offset`     = offset,
      `$$app_token` = app_token
    ) |>
    req_user_agent("STA9750-MP03-EC2/Arati-K") |>
    req_error(is_error = function(resp) resp_status(resp) >= 400)

  resp <- req_perform(req)
  writeBin(resp_body_raw(resp), out)
  Sys.sleep(pause)
  out
}

socrata_download_all <- function(url,
                                 fmt       = c("csv","json"),
                                 limit     = 50000L,
                                 max_pages = Inf,
                                 dir_out   = ec_dir,
                                 pause     = 0.4) {
  fmt <- match.arg(fmt)
  files  <- character()
  offset <- 0L
  page   <- 1L
  repeat {
    if (page > max_pages) break
    f <- socrata_download_page(url, limit, offset, fmt, dir_out, pause)
    files <- c(files, f)

    # rows returned on this page
    nret <- if (fmt == "csv") {
      nrow(read_csv(f, show_col_types = FALSE, progress = FALSE))
    } else {
      nrow(as_tibble(fromJSON(f, simplifyDataFrame = TRUE)))
    }

    if (nret < limit) break
    offset <- offset + limit
    page   <- page   + 1L
  }
  unique(files)
}

read_bind_paged <- function(files, fmt = c("csv", "json")) {
  fmt <- match.arg(fmt)
  if (fmt == "csv") {
    files |>
      lapply(function(f) read_csv(f, show_col_types = FALSE, progress = FALSE)) |>
      bind_rows()
  } else {
    files |>
      lapply(function(f) as_tibble(fromJSON(f, simplifyDataFrame = TRUE))) |>
      bind_rows()
  }
}
```

```{r}
#| label: ec2-download-datasets
# Forestry Risk Assessments
risk_url <- "https://data.cityofnewyork.us/resource/259a-b6s7.json"
risk_files <- socrata_download_all(risk_url, fmt = "json", limit = 50000)
risk_raw   <- read_bind_paged(risk_files, fmt = "json")

# Forestry Work Orders
wo_url   <- "https://data.cityofnewyork.us/resource/bdjm-n7q4.json"
wo_files <- socrata_download_all(wo_url,   fmt = "json", limit = 50000)
wo_raw   <- read_bind_paged(wo_files, fmt = "json")

# quick sanity prints (optional)
nrow(risk_raw); length(names(risk_raw))
nrow(wo_raw);   length(names(wo_raw))
```
